Sorting by Reversals

### 2020-06-29 slides.tex
improvements

\usepackage{biblatex}
\addbibresource{sources.bib}

@@

\begin{algorithm}[Watterson]
For each element at position $i$, perform $\rho(i, \pi_i^{-1})$ if $i \neq \pi_i$. \cite{Kececioglu1995}
\end{algorithm}\pause\padding

@@

An alternating cycle in $G(\pi)$ is a \b{$k$-cycle} if it has $k$ constituting red edges.
\end{definition}

\begin{definition}
We call an alternating cycle $C$ in $G(\pi)$ \b{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges of $C$. \pause\spadding

An alternating cycle in $G(\pi)$ is \b{directed} with respect to $\pi$ if it is possible to walk along the whole cycle traversing each red edge in the direction of its orientation.
\end{definition}

\begin{proposition}[\citeauthor*{Caprara1997}]
An alternating cycle in $G(\pi)$ is oriented iff it is undirected.
\end{proposition}\pause

\begin{exampleblock}{Example}
\begin{center}
\begin{tikzpicture}
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (6) at (-3.75, 0) {0};
		\node [style=node] (7) at (-2.25, 0) {3};
		\node [style=node] (8) at (-0.75, 0) {4};
		\node [style=node] (9) at (0.75, 0) {1};
		\node [style=node] (10) at (2.25, 0) {2};
		\node [style=node] (11) at (3.75, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (6) to (7);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (10) to (11);
		\draw [style=blue edge, bend right] (6) to (9);
		\draw [style=blue edge, bend left] (7) to (10);
		\draw [style=blue edge, bend right] (8) to (11);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}\pause
$\implies$ directed and unoriented.
\end{exampleblock}

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$. Given an alternating cycle decomposition $C$ of $G(\pi)$ construct $R(C)$.\spadding

\begin{tabularx}{\textwidth}{XR}
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture}
with
{\begin{align*}
    C = \{&\{\{0,4\},\{3,4\},\{1,3\},\{0,1\}\} \\
          &\{\{2,4\},\{2,3\},\{3,5\},\{4,5\}\}\}
\end{align*}} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=red node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=blue node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=blue node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (1) to (0);
		\draw (2) to (1);
		\draw (0) to (3);
		\draw (1) to (3);
		\draw (2) to (0);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}
\end{exampleblock}\pause

@@

\begin{tabularx}{\textwidth}{X|R}
$G(\pi)$ and $R(C)$ & $G(\pi \circ \rho(u))$ and $R(C_u)$ \\ \hline
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {1};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {4};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (4) to (5);
		\draw [style=red edge] (8) to (9);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (5) to (9);
	\end{pgfonlayer}
\end{tikzpicture}\\
\spadding\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=red node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=blue node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=blue node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (1) to (0);
		\draw (2) to (1);
		\draw (0) to (3);
		\draw (1) to (3);
		\draw (2) to (0);
	\end{pgfonlayer}
\end{tikzpicture} &
\spadding\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=blue node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=blue node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=red node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=red node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (2) to (3);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}
\end{exampleblock}

>>>>

\usepackage[sorting=ynt]{biblatex}
\addbibresource{sources.bib}

@@

\begin{algorithm}[ratchet algorithm, \citeauthor*{Watterson19821}]
For each element at position $i$, perform $\rho(i, \pi_i^{-1})$ if $i \neq \pi_i$.
\end{algorithm}\pause

@@

An alternating cycle in $G(\pi)$ is a \b{$k$-cycle} if it has $k$ constituting red edges. \pause\spadding

We call an alternating cycle $C$ in $G(\pi)$ \b{oriented} if there is a $1$- or $2$-reversal acting on two constituting red edges of $C$.
\end{definition}

@@

\begin{exampleblock}{Example}
Let $\pi = (0\ 4\ 2\ 1\ 3\ 5) \in S_4$. \par
Given the alternating cycle decomposition $C$ of $G(\pi)$

\begin{tabularx}{\textwidth}{XR}
{\footnotesize\begin{align*}
    C = \{&\{\{0,4\},\{3,4\},\{1,3\},\{0,1\}\} \\
          &\{\{2,4\},\{2,3\},\{3,5\},\{4,5\}\}\}
\end{align*}} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}\pause

construct $R(C)$.

\begin{center}
\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=red node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=blue node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=blue node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (1) to (0);
		\draw (2) to (1);
		\draw (0) to (3);
		\draw (1) to (3);
		\draw (2) to (0);
	\end{pgfonlayer}
\end{tikzpicture}
\end{center}
\end{exampleblock}\pause

@@

\begin{tabularx}{\textwidth}{X|R}
$G(\pi)$ and $R(C)$ & $G(\pi \circ \rho(u))$ and $R(C_u)$ \\ \hline
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {4};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {1};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (3) to (4);
		\draw [style=red edge] (8) to (9);
		\draw [style=red edge] (0) to (3);
		\draw [style=red edge] (5) to (8);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (3) to (9);
		\draw [style=blue edge, bend right] (0) to (5);
		\draw [style=blue edge, bend right] (3) to (8);
	\end{pgfonlayer}
\end{tikzpicture} &
\begin{tikzpicture}[scale=0.6, every node/.style={scale=0.6}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=node] (0) at (-3, 0) {0};
		\node [style=node] (3) at (-1.5, 0) {1};
		\node [style=node] (4) at (0, 0) {2};
		\node [style=node] (5) at (1.5, 0) {4};
		\node [style=node] (8) at (3, 0) {3};
		\node [style=node] (9) at (4.5, 0) {5};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw [style=red edge] (4) to (5);
		\draw [style=red edge] (8) to (9);
		\draw [style=blue edge, bend left] (4) to (8);
		\draw [style=blue edge, bend left] (5) to (9);
	\end{pgfonlayer}
\end{tikzpicture}\\
\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=red node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=red node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=blue node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=blue node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (1) to (0);
		\draw (2) to (1);
		\draw (0) to (3);
		\draw (1) to (3);
		\draw (2) to (0);
	\end{pgfonlayer}
\end{tikzpicture} &
\spadding\begin{tikzpicture}[scale=0.8, every node/.style={scale=0.8}]
	\begin{pgfonlayer}{nodelayer}
		\node [style=blue node] (0) at (0, 1) {$\{3,4\}$};
		\node [style=blue node] (1) at (-2, 1) {$\{0,1\}$};
		\node [style=red node] (2) at (-2, -1) {$\{2,3\}$};
		\node [style=red node] (3) at (0, -1) {$\{4,5\}$};
		\node [style=blue node] (4) at (2, 0) {$\{1,2\}$};
	\end{pgfonlayer}
	\begin{pgfonlayer}{edgelayer}
		\draw (2) to (3);
	\end{pgfonlayer}
\end{tikzpicture}
\end{tabularx}
\end{exampleblock}
